---
title: "CIA Factbook Visualizations"
author: "Cecily Hansen"
date: "2026-03-10"
format: html
---

## Introduction

I am working with data from the CIA World Factbook, sourced brom TidyTuesday at this url: https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-10-22/readme.md. The World Factbook is a website run by the U.S. Central Intelligence Agency that provides information on the people, government, economy, and many other variables for all countries. The TidyTuesday data set I utilized has data from 2014. When I clicked on the link provided by the Github page, I was redirected to a page from the CIA website that informed me that the World Factbook is no longer being updated or maintained as of February 2026. When I researched why, the general consensus was that the resource is outdated in the "digital age". Since the public domain is no longer available, GitHub repositories like the one I utilized are now the only homes to the archived data from the World Factbook. My data set has 259 observations of 11 variables. I am most interested in looking at population and population growth rate. I also looked at infant mortality rate, which is measured as the number of deaths of infants under one year old per 1,000 live births. I am interested to see what countries have the highest populations, and how different countries populations are changing.

## Primary Visualizations

```{r}
library(tidyverse)
library(maps)
library(ggplot2)


cia_factbook <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-10-22/cia_factbook.csv') |>
  mutate(country = if_else(country == "United States", true = "USA", false = country))
```

### Visualization 1 - Global Population

I chose to make a Mercator projection with the fill set to population. I joined the CIA Factbook data to map data, which allowed me to create the map fairly easily. The map shows that China and India have the largest populations by far, with the USA, Russia, Brazil, and Indonesia behind. This matches with my prior expectation of global populations.

```{r}
world_map <- map_data("world")

world_joined <- left_join(world_map, cia_factbook, join_by(region == country))

ggplot(data = world_joined, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = population)) +
  coord_map(projection = "mercator", xlim = c(-180,180)) +
  scale_fill_distiller(type = "seq", palette = "Blues", direction = 1) +
  theme_minimal() +
  labs(title = "Global Population", caption = "Data sourced from the CIA World Factbook 2014")
```

### Visualization 2 - Infant Mortality

For this visualization, I used a Gilbert projection to represent infant mortality rates in each country. This variable is measured by the number of deaths of infants under one year per 1,000 live births. The map shows that higher mortality rates are concentrated more heavily in the Global South, particularly in Africa. Afghanistan has the highest infant mortality rate of 117.23 deaths for every 1,000 babies born. Visualizations like these highlight inequalities in healthcare. Western countries have higher access to crucial healthcare, while developing nations suffer. This map highlights area where financial resources are necessary to improve mother and infant health and welfare.

```{r}
ggplot(data = world_joined, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = infant_mortality_rate)) +
  coord_map(projection = "gilbert", xlim = c(-180,180)) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Global Infant Mortality", caption = "Data sourced from the CIA World Factbook 2014")
```

### Visualization 3 - Population Change

For my final visualization, I created bar plots representing population growth in a randomly selected sample of countries. I was originally wanting to include every country, but that did not fit well onto the plot. I used slice_sample() to pull 30 random countries/territories twice to make two visualizations. The countries are ordered by population growth rate, with the highest growth rates at the top and lowest at the bottom. In the first visualization, the bottom 7 countries have negative growth rates, meaning that their population is shrinking. When thinking about the regions with the highest growth rates, it's mainly African and Middle Eastern countries. European countries typically have low or negative growth rates. African countries have high birth rates because of the need for child labor, lack of access to contraception, low female education levels, and early marriage. Western countries have much lower growth rates because labor is outsourced, there is better education and healthcare, and having less children is socially acceptable. The second plot shows a similar trend, where Western and developed countries have lower population growth than undeveloped countries.

```{r}
factbook_slice <- cia_factbook |> slice_sample(n = 30) |>
  select(country, population_growth_rate) |>
  arrange(population_growth_rate) |>
  filter(!is.na(population_growth_rate))

ggplot(data = factbook_slice, aes(x = reorder(country, population_growth_rate), y = population_growth_rate)) +
  geom_bar(stat = "identity", fill = "darkred") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Population Growth Rate in 2014 (Plot 1)", x = "Country", y = "Population Growth Rate", caption = "Data sourced from the CIA World Factbook 2014")

factbook_slice2 <- cia_factbook |> slice_sample(n = 30) |>
  select(country, population_growth_rate) |>
  arrange(population_growth_rate) |>
  filter(!is.na(population_growth_rate))

ggplot(data = factbook_slice2, aes(x = reorder(country, population_growth_rate), y = population_growth_rate)) +
  geom_bar(stat = "identity", fill = "darkred") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Population Growth Rate in 2014 (Plot 2)", x = "Country", y = "Population Growth Rate", caption = "Data sourced from the CIA World Factbook 2014")
```

## Conclusion

If I had more time to work with this data, I would like to examine the internet_users variable. It would be interesting to look at how internet use varies from country to country. I would also have liked to look at a more recent set of data from 2025 or 2026. I could join the data sets together and make faceted maps to see how certain variables have changed in the past decade. It could also be interesting to make a linear model with some of the variables and have population growth rate be the response variable. I could examine what variables have the the most impact - for example, would we expect countries with a high migration rate to have a higher population? I am not sure how well a linear model would work with this data set but it would be interesting to explore. My visualizations connect to class because I used ideas we discuss to make my plots. I made several maps utilizing the maps package and geom_polygon(). I used two different projections to visualize the globe, and I used the appropriate color palettes for sequential data. I also changed themes and added titles to make my plots easy to understand. I believe that my plots accurately represent the data I am sharing.
